<<<<<<< HEAD
---
title: "Exploration of processed ChIP-seq data"
author: "Kaisu Hiltunen"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(rtracklayer)
  library(epiwraps)
  library(GenomicRanges)
  library(R.utils)
})
ah <- AnnotationHub()
```

# Assignment week4

Find and download from ENCODE the peaks (i.e. bed-like format) for the following in mouse embryonic stem cells (mESC) :

-   Transcription factor p300

    histone modifications

-   H3K4me3,

-   H3K4me1,

-   H3K27ac, and

-   H3K27me3

# Downloading the processed data

```{r, eval=FALSE}
dir.create("Assignment_data")

#Downloading the H3K27me3 data:
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz","Assignment_data/H3K27me3.bed.gz")
#gunzip("Assignment_data/H3K27me3.bed.gz")

#Downloading the H3K27ac data:
download.file("https://www.encodeproject.org/files/ENCFF360VIS/@@download/ENCFF360VIS.bed.gz",
 "Assignment_data/H3K27ac.bed.gz")
#gunzip("Assignment_data/H3K27ac.bed.gz")

#Downloading the H3K4me1 data: 
download.file("https://www.encodeproject.org/files/ENCFF333IJH/@@download/ENCFF333IJH.bed.gz",
 "Assignment_data/H3K4me1.bed.gz")
#gunzip("Assignment_data/H3K4me1.bed.gz")

#Downloading the H3K4me3 data:
download.file("https://www.encodeproject.org/files/ENCFF247GVM/@@download/ENCFF247GVM.bed.gz",
 "Assignment_data/H3K4me3.bed.gz")
#gunzip("Assignment_data/H3K4me3.bed.gz")

#p300
download.file("https://www.encodeproject.org/files/ENCFF653YVR/@@download/ENCFF653YVR.bed.gz", "Assignment_data/p300.bed.gz")
```

# specifying the format

Importing and setting the format to narrowpeak

```{r}
#p300
peaks_p300 <- rtracklayer::import("Assignment_data/p300.bed.gz", format="narrowPeak")

#H3K4me1
peaks_H3K4me1 <- rtracklayer::import("Assignment_data/H3K4me1.bed.gz", format="narrowPeak")

#H3K4me3
peaks_H3K4me3 <- rtracklayer::import("Assignment_data/H3K4me3.bed.gz", format="narrowPeak")

#H3K27ac
peaks_H3K27ac <- rtracklayer::import("Assignment_data/H3K27ac.bed.gz", format="narrowPeak")

#H3K27me3
peaks_H3K27me3 <- rtracklayer::import("Assignment_data/H3K27me3.bed.gz", format="narrowPeak")


head(peaks_H3K27me3)
```

```{r, eval=FALSE,include=FALSE}
#This is possibly redundant for this task. Might delete later.
q <- query(ah, c("mm10","ensembl"))
ensembl_mm10<-ah[["AH98194"]]
ensembl_mm10
mm10<-ah[["AH6066"]]
mm10
```

Extract the subset of peaks which overlap with p300.

```{r, include=FALSE, eval=FALSE}
#p300 overlap H3K4me1
peaks_p300$overlapsH3K4me1<-overlapsAny(peaks_p300,peaks_H3K4me1)

#p300 overlap H3K4me3
peaks_p300$overlapsH3K4me3<-overlapsAny(peaks_p300,peaks_H3K4me3)

#p300 overlap H3K27ac
peaks_p300$overlapsH3K27ac<-overlapsAny(peaks_p300,peaks_H3K27ac)

#p300 overlap H3K27me3
peaks_p300$overlapsH3K27me3<-overlapsAny(peaks_p300,peaks_H3K27me3)

```
```{r}
proportion=c()
absolute_overlaps=c()
histone=c()
histone=c('H3K4me1','H3K4me3','H3K27ac','H3K27me3')

#p300 overlap H3K4me1
peaks_H3K4me1$overlapsp300<-overlapsAny(peaks_H3K4me1,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K4me1$overlapsp300)/length(peaks_H3K4me1))
absolute_overlaps<-append(absolute_overlaps, sum(peaks_H3K4me1$overlapsp300))


#p300 overlap H3K4me3
peaks_H3K4me3$overlapsp300<-overlapsAny(peaks_H3K4me3,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K4me3$overlapsp300)/length(peaks_H3K4me3))
absolute_overlaps<-append(absolute_overlaps, sum(peaks_H3K4me3$overlapsp300))

#p300 overlap H3K27ac
peaks_H3K27ac$overlapsp300<-overlapsAny(peaks_H3K27ac,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K27ac$overlapsp300)/length(peaks_H3K27ac))
absolute_overlaps<-append(absolute_overlaps, sum(peaks_H3K27ac$overlapsp300))


#p300 overlap H3K27me3
peaks_H3K27me3$overlapsp300<-overlapsAny(peaks_H3K27me3,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K27me3$overlapsp300)/length(peaks_H3K27me3))
absolute_overlaps<-append(absolute_overlaps, sum(peaks_H3K27me3$overlapsp300))


```

```{r}
df <- data.frame(histone, absolute_overlaps, proportion)

df
```

```{r}
print(paste0("Total amount of marks overlapped by p300: ", sum(df$absolute_overlaps)))
```

=======
---
title: "Exploration of processed ChIP-seq data"
author: "Kaisu Hiltunen"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(rtracklayer)
  library(epiwraps)
  library(GenomicRanges)
  library(R.utils)
})
ah <- AnnotationHub()
```

# Assignment week4

Find and download from ENCODE the peaks (i.e. bed-like format) for the following in mouse embryonic stem cells (mESC) :

-   Transcription factor p300

    histone modifications

-   H3K4me3,

-   H3K4me1,

-   H3K27ac, and

-   H3K27me3

# Downloading the processed data

```{r, eval=FALSE}
dir.create("Assignment_data")

#Downloading the H3K27me3 data:
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz","Assignment_data/H3K27me3.bed.gz")
#gunzip("Assignment_data/H3K27me3.bed.gz")

#Downloading the H3K27ac data:
download.file("https://www.encodeproject.org/files/ENCFF360VIS/@@download/ENCFF360VIS.bed.gz",
 "Assignment_data/H3K27ac.bed.gz")
#gunzip("Assignment_data/H3K27ac.bed.gz")

#Downloading the H3K4me1 data: 
download.file("https://www.encodeproject.org/files/ENCFF333IJH/@@download/ENCFF333IJH.bed.gz",
 "Assignment_data/H3K4me1.bed.gz")
#gunzip("Assignment_data/H3K4me1.bed.gz")

#Downloading the H3K4me3 data:
download.file("https://www.encodeproject.org/files/ENCFF247GVM/@@download/ENCFF247GVM.bed.gz",
 "Assignment_data/H3K4me3.bed.gz")
#gunzip("Assignment_data/H3K4me3.bed.gz")

#p300
download.file("https://www.encodeproject.org/files/ENCFF653YVR/@@download/ENCFF653YVR.bed.gz", "Assignment_data/p300.bed.gz")
```

# specifying the format

Importing and setting the format to narrowpeak

```{r}
#p300
peaks_p300 <- rtracklayer::import("Assignment_data/p300.bed.gz", format="narrowPeak")

#H3K4me1
peaks_H3K4me1 <- rtracklayer::import("Assignment_data/H3K4me1.bed.gz", format="narrowPeak")

#H3K4me3
peaks_H3K4me3 <- rtracklayer::import("Assignment_data/H3K4me3.bed.gz", format="narrowPeak")

#H3K27ac
peaks_H3K27ac <- rtracklayer::import("Assignment_data/H3K27ac.bed.gz", format="narrowPeak")

#H3K27me3
peaks_H3K27me3 <- rtracklayer::import("Assignment_data/H3K27me3.bed.gz", format="narrowPeak")


head(peaks_H3K27me3)
```

```{r, eval=FALSE,include=FALSE}
#This is possibly redundant for this task. Might delete later.
q <- query(ah, c("mm10","ensembl"))
ensembl_mm10<-ah[["AH98194"]]
ensembl_mm10
mm10<-ah[["AH6066"]]
mm10
```

Extract the subset of peaks which overlap with p300.

```{r, include=FALSE, eval=FALSE}
#p300 overlap H3K4me1
peaks_p300$overlapsH3K4me1<-overlapsAny(peaks_p300,peaks_H3K4me1)

#p300 overlap H3K4me3
peaks_p300$overlapsH3K4me3<-overlapsAny(peaks_p300,peaks_H3K4me3)

#p300 overlap H3K27ac
peaks_p300$overlapsH3K27ac<-overlapsAny(peaks_p300,peaks_H3K27ac)

#p300 overlap H3K27me3
peaks_p300$overlapsH3K27me3<-overlapsAny(peaks_p300,peaks_H3K27me3)

```
```{r}
proportion=c()
absolute_overlaps=c()
histone=c()
histone=c('H3K4me1','H3K4me3','H3K27ac','H3K27me3')

#p300 overlap H3K4me1
peaks_H3K4me1$overlapsp300<-overlapsAny(peaks_H3K4me1,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K4me1$overlapsp300)/length(peaks_H3K4me1))
absolute_overlaps=append(absolute_overlaps, sum(peaks_H3K4me1$overlapsp300))


#p300 overlap H3K4me3
peaks_H3K4me3$overlapsp300<-overlapsAny(peaks_H3K4me3,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K4me3$overlapsp300)/length(peaks_H3K4me3))
absolute_overlaps=append(absolute_overlaps, sum(peaks_H3K4me3$overlapsp300))

#p300 overlap H3K27ac
peaks_H3K27ac$overlapsp300<-overlapsAny(peaks_H3K27ac,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K27ac$overlapsp300)/length(peaks_H3K27ac))
absolute_overlaps=append(absolute_overlaps, sum(peaks_H3K27ac$overlapsp300))


#p300 overlap H3K27me3
peaks_H3K27me3$overlapsp300<-overlapsAny(peaks_H3K27me3,peaks_p300)
proportion<-append(proportion, sum(peaks_H3K27me3$overlapsp300)/length(peaks_H3K27me3))
absolute_overlaps=append(absolute_overlaps, sum(peaks_H3K27me3$overlapsp300))


```

```{r}
df <- data.frame(histone, absolute_overlaps, proportion)

df
```

```{r}
print(paste0("Total amount of marks overlapped by p300: ", sum(df$absolute_overlaps)))
```

>>>>>>> 823faccbd1669b3c966430c4f6f749f371e96f0e
